Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-14 talk/pepm2011/paper.tex
fix an XXX, add a new one

The core of our trace optimization technique can be
viewed as partial evaluation: the partial evaluation
performs a form of escape analysis \cite{bruno_blanchet_escape_2003} on the traces and make some
objects that are allocated in the trace \emph{static}\footnote{These objects are
called \emph{virtual} in Psyco \cite{rigo_representation-based_2004}.} which
means that they do not occur any more in the optimized trace. This technique is
informally described in Section~\ref{sec:statics}, a more formal description is
given in Section~\ref{sec:formal}.

@@

XXX add a footnote saying that "static" is meant in the sense of PE, not static
allocation

@@

The partial evaluation works by traversing the trace from beginning to end.
Whenever a \lstinline{new} operation is seen, the operation is removed and a static
object is constructed and associated with the variable that would have stored
the result of \lstinline{new}. The static object describes the shape of the
original object, \eg where the values that would be stored in the fields of the
allocated object come from, as well as the type of the object. Whenever the
optimizer sees a \lstinline{set} that writes into such an object, that shape
description is updated and the operation can be removed, which means that the
operation was done at partial evaluation time. When the optimizer encounters a
\lstinline{get} from such an object, the result is read from the shape
description, and the operation is also removed. Equivalently, a
\lstinline{guard_class} on a variable that has a shape description can be removed
as well, because the shape description stores the type and thus the outcome of
the type check the guard does is statically known.

@@

>>>>

The core of our trace optimization technique can be
viewed as partial evaluation: the partial evaluation
performs a form of escape analysis \cite{bruno_blanchet_escape_2003} on the traces and make some
objects that are allocated in the trace \emph{static} which
means that they do not occur any more in the optimized trace. This technique is
informally described in Section~\ref{sec:statics}, a more formal description is
given in Section~\ref{sec:formal}.

@@

@@

The partial evaluation works by traversing the trace from beginning to end.
Whenever a \lstinline{new} operation is seen, the operation is removed and a \empth{static
object}\footnote{Here ``static'' is meant in the sense of partial
evaluation, \ie known at partial evaluation time, not in the sense of static
allocation or static method.} is constructed and associated with the variable
that would have stored
the result of \lstinline{new}. The static object describes the shape of the
original object, \eg where the values that would be stored in the fields of the
allocated object come from, as well as the type of the object. Whenever the
optimizer sees a \lstinline{set} that writes into such an object, that shape
description is updated and the operation can be removed, which means that the
operation was done at partial evaluation time. When the optimizer encounters a
\lstinline{get} from such an object, the result is read from the shape
description, and the operation is also removed. Equivalently, a
\lstinline{guard_class} on a variable that has a shape description can be removed
as well, because the shape description stores the type and thus the outcome of
the type check the guard does is statically known.

@@

XXX Psyco \cite{rigo_representation-based_2004}

XXX cite SELF paper for the type propagation effects
